---
title: "ILC008"
author: "Rebecca Payne"
date: "23/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(readxl) 
library(CATALYST)
library(cowplot)
library(flowCore)
library(scater)
library(SingleCellExperiment)
library(openxlsx)
library(markdown)
```


### Create SCE and use FLOWSom to identify monocytes
Load ILC008 FCS files (CD45+), metadata, panel and create flowset

```{r}
md <- "metadata_matched_v3.xlsx" 
md <- read_excel(md)       
head(data.frame(md)) 
```


```{r}
fs <- read.flowSet(md$file_name, transformation = F, truncate_max_range = F)
```

```{r}
panel <- "Panel_matched_v3.xlsx" 
panel <- read_excel(panel)                                         
head(data.frame(panel)) 
```


```{r}
all(panel$fcs_colname %in% colnames(fs))
```

```{r}
setdiff(fs@frames$ilc008_BMT011_200619_15_normalized_CD45.fcs@parameters@data$desc,panel$antigen)
```


```{r}
md$condition <- factor(md$condition, levels = c("none"))      
md$date <- factor(md$timepoint, levels = c("none"))
md$patient_id <- factor(md$patient_id, levels = c("ILC008", "HC002"))

md$sample_id <- factor(md$sample_id, levels = md$sample_id[order(md$sample_id)])  
```

```{r}
sce <- prepData(fs, panel, md, features = panel$fcs_colname, md_cols = list(file = "file_name", id = "sample_id", 
                                                                            factors = c("condition", "timepoint", "patient_id")))

```


## Overview analysis of Cytof data

```{r}
n_cells(sce)
```

```{r}
plotCounts(sce, color_by = "sample_id")
```

```{r}
plotNRS(sce, features = type_markers(sce), color_by = "sample_id")
```


## Identify monocytes for removal 

Cluster FLOWSom analysis - max 20 clusters

For reproducibility set random seed

```{r}
set.seed(4321)
sce <- cluster(sce, features = "type", 
               xdim = 10, ydim = 10, maxK = 20, 
               verbose = FALSE, seed = 4321) 
```

## Dimension reduction - TSNE
```{r}
set.seed(4321)
sce <- runDR(sce, dr = "TSNE", cells = 500, features = "type")
```

```{r}
plotDR(sce, "TSNE", color_by = "meta15")
```


```{r}
plotDR(sce, "TSNE", color_by = "meta15") + facet_wrap("patient_id") +
  guides(color = guide_legend(ncol = 2, override.aes = list(size = 3))) 
```


```{r fig.width=10}
plotMultiHeatmap(sce, 
    k = "meta15", scale = "never",
    hm1 = FALSE, hm2 = c("CD14", "CD16", "CD56", "CD4"),
    row_anno = FALSE, col_clust = FALSE,
    hm2_pal = c("grey95", "black"))
```

### Monocytes identified by CD14+, CD16+ and CD4low expression - cluster 1

### Create new sce (sce1) with monocyte clustes removed

```{r}
sce1 <- filterSCE(sce, cluster_id %in% c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15),  k = "meta15")
```

### FlowSOM cluster of sce1
```{r}
set.seed(4321)
sce1 <- cluster(sce1, features = "type", 
               xdim = 10, ydim = 10, maxK = 20, 
               verbose = FALSE, seed = 4321) 
```

### TSNE Dimension reduction of sce1
```{r}
set.seed(4321)                                                                      
sce1 <- runDR(sce1, dr = "TSNE", cells = 500, features = "type")
```

```{r}
plotDR(sce1, "TSNE", color_by = "meta20") + facet_wrap("sample_id") +
  guides(color = guide_legend(ncol = 2, override.aes = list(size = 3)))   
```

```{r}
plotDR(sce1, "TSNE", color_by = "meta20")
```



```{r}
plotExprHeatmap(sce1, features = "type",
    by = "cluster_id", k = "meta20", m = "meta8",
    scale = "last", q = 0.01, perc = TRUE, col_dend = FALSE)
```

```{r fig.width=15}
plotPbExprs(sce1, k = "meta20", features = "type", 
  group_by = "cluster_id", color_by = "sample_id", 
  size_by = TRUE, geom = "points", jitter = FALSE, ncol = 5)
```


```{r fig.width=10}
plotMultiHeatmap(sce1, 
    k = "meta20", scale = "never",
    hm1 = FALSE, hm2 = c("CD45RA", "CD16", "CD56"),
    row_anno = FALSE, col_clust = FALSE,
    hm2_pal = c("grey95", "black"))
```

Cluster 3, 11 : NK_cells


```{r fig.width=20}
plotMultiHeatmap(sce1, 
    k = "meta20", scale = "never",
    hm1 = FALSE, hm2 = c("CD3", "CD4", "CD8", "TCRgd", "GranB", "CD45RA", "CD27", "CD28", "CD127", "FOXP3", "CD25", "Valpha7.2"),
    row_anno = FALSE, col_clust = FALSE,
    hm2_pal = c("grey95", "black"))
```

CLuster 7 : TCRgd

Cluster 15, 20 : CD4_CM
Cluster 12 :CD4_naive
Cluster 4 : CD8_naive
Cluster 9, 8 : CD3_MAIT
Cluster 6 : CD8_Cytotoxic

```{r}
plotDR(sce1, "TSNE", color_by = c("CD4", "FOXP3", "CD127", "CD25"))
```



```{r}
plotDR(sce1, "TSNE", color_by = ("meta20"))
```



```{r fig.width=15}
plotMultiHeatmap(sce1, 
    k = "meta20", scale = "never",
    hm1 = FALSE, hm2 = c("CD19", "CD20", "CD3", "CD4", "IgD", "CD27"),
    row_anno = FALSE, col_clust = FALSE,
    hm2_pal = c("grey95", "black"))
```

Cluster 1, 2, : B_naive
Cluster 10, 5 : B_memory

```{r}
plotDR(sce1, "TSNE", color_by = c("HLADR", "CD14", "CD16", "S100a9"))
```


```{r fig.width=20}
plotMultiHeatmap(sce1, 
    k = "meta20", scale = "never",
    hm1 = FALSE, hm2 = c("CD123", "SIGLEC6", "AXL", "GranB", "CD1c", "FCERI", "CD11c", "HLADR", "S100a9", "CD14", "CD16", "CLEC9A", "CD49d"),
    row_anno = FALSE, col_clust = FALSE,
    hm2_pal = c("grey95", "black"))
```

cluster 14  : DC6(pDC)_1
cluster 17.18: Neutrophil
cluster 13 :DC4
cluster 16 : Eosinophils
cluster 19: Basophils




## Re-label meta20 clusters


```{r}
merging_table1 <- "merging_v3.xlsx"   
merging_table1 <- read_excel(merging_table1)          
head(data.frame(merging_table1)) 
```

```{r}
merging_table1$new_cluster <- factor(merging_table1$new_cluster,         
                                     levels = c("B_naive", "NK_cells", "CD8_naive", "CD8_cytotoxic", "CD3_TCRgd", "CD3_MAIT", "B_memory", "CD4_naive", "DC4", "DC6(pDC)", "CD4_CM", "Eosinophil", "Neutrophil", "Basophil" )) 
```

```{r}
sce1 <- mergeClusters(sce1, k = "meta20",                                  
                     table = merging_table1, id = "merging1") 
```


```{r}

plotAbundances(sce1, k = "merging1", by = "sample_id", group_by = "condition")

```

```{r}
newSCEILC008 <- filterSCE(sce1, patient_id == "ILC008")

```


```{r}

plotAbundances(newSCEILC008, k = "merging1", by = "sample_id", group_by = "condition")

```

```{r}
pA <- plotAbundances(newSCEILC008, k = "merging1", by = "sample_id")
pA$data
```

## Export frequencies of merged cluster to dataframe

```{r}
ns <- table(
  cluster_id = cluster_ids(newSCEILC008, k= "merging1"), 
  sample_id = sample_ids(newSCEILC008))
fq <- prop.table(ns, 2) * 100
df <- as.data.frame(fq)
```

```{r}
getwd()
```

```{r}
write.xlsx (df,  "/Users/rebeccapayne/OneDrive - Newcastle University/ILC_study/Cytof_FCS_files/WBC/20190912_ILC008_WBC/RawFCSfiles/normed/ilc008_norm_singlets_live_CD45/Git_ILC008/ILC008_Git/mydata.xlsx", row.names = TRUE)

```


```{r}
n_cells <- table(sample = sce1$sample_id, cluster = cluster_ids(sce1, "merging1"))

write.xlsx(n_cells, "n_cells.xlsx")
```



```{r}
sessionInfo()
```






